\textbf{ Interpretable Gaussian Processes for Stellar Variability\\ I: Static Surfaces }

### 2020-12-23 tex/ms.tex
more figures, more text

\xxx{TODO: discuss the batch calibration default run.}
Stellar inclination residuals (posterior mean minus true value) normalized to the
posterior standard deviation. Only \emph{marginally} overestimate the variance.
Tiny sunspots are about $70{,}000$ km across, comparable to large sunspots.

@@

@@

\xxx{More runs: 1) Unnormalized (to show what we can learn if we know the baseline, as in
    multi-band photometry. 2) Big spots. 3) Single spot. 4) Variance in everything.)}

@@

\subsection{Miscellaneous}
\label{sec:integerspots}
Shrinkage prior on sigma. Inclination distribution is bimodal.

@@

>>>>

The results in
Figures~\ref{fig:calibration_default_corner}---\ref{fig:calibration_default_inclination}
are based on a single realization of the light curve ensemble conditioned on
the properties of Table~\ref{tab:synthetic}. To properly gauge potential biases
in our model, it is useful to re-run the experiment under many different realizations
of the dataset. We therefore generate 100 ensembles of light curves in
exactly the same way as above and run inference on each of them.
Figure~\ref{fig:calibration_batch_default_corner} shows the marginal and joint
posterior distributions for $\pmb{\theta}_\bullet$ for all 100 trials.
Posteriors for individual trials are shown as the transluscent colored curves
(in the marginal plots) and as ellipses bounding the $1\sigma$ posterior level
(in the joint posterior plots). The black curves show the marginal distributions of all
samples across all trials, and the black contours show the corresponding $1\sigma$
levels in the joint posterior plots.

If our model is truly unbiased, in the limit of an infinite number of realizations
of the data, the expectation value of the distribution
of samples across all ensembles (the mean of the black curves in the marginal
posterior plots) should coincide with the true values (blue lines).
This is the case with the spot size $r$ posterior: on average, the posterior
distributions are centered on the correct value.
However, this is not the case for
the spot latitude parameters $\mu_\phi$ and $\sigma_\phi$, for which our
posterior means are biased high. While the \emph{modes} of their
posteriors are very close to the true values, the distributions have long
tails extending to high latitudes and high variance, respectively.

The reason for this bias has to do with the baseline degeneracy discussed in
\S\ref{sec:baseline} and illustrated in Figure~\ref{fig:mean_normalization}:
the total spottiness of a stellar surface cannot be probed in relative
photometry. In particular, this means spots near the poles lie almost
entirely in the null space. Applied to the problem at hand, this degeneracy
makes it difficult to distinguish between stars with spots concentrated
exclusively at mid-latitudes (in this case, the truth) and
stars with spots centered closer to the poles but with large latitudinal
variance. The latter configuration leads to many spots close to the poles,
whose effect on the (relative) light curve is negligible, and some spots
at mid-latitudes, whose effect on the light curve is similar to that of
the former configuration.
Thus, the data alone cannot be used to discriminate between these two
scenarios, introducing the degeneracy we see in the posterior. In fact,
it is clear that in the tails of the distribution, the mean spot latitude
and the standard deviation of spot latitudes are positively correlated.
The bias we see is therefore not a shortcoming of the model, but of the
\emph{data} itself. To get around this, we either need to impose a
shrinkage prior on $\sigma_\phi$ (\S\ref{sec:sigma-prior}),
observe in a different mode that is sensitive to baseline information
(\S\ref{sec:breaking_baseline_degeneracy}), or simply collect more data.
As we will see below, the degeneracy described above is not perfect:
for very large ($M \sim 1{,}000$) ensembles, high-variance polar spots
can confidently be ruled out.

Lastly, the posteriors for the contrast $c$ and the number of spots $n$
are mostly unbiased. The contrast distribution has a bit of
a tail; inspection of Figure~\ref{fig:calibration_batch_default_corner}
reveals that it too is positively correlated with the mean spot latitude
and therefore suffers from the same degeneracy as above. And while the
mean of the spot number distribution is roughly correct, the posterior
is nearly unchanged across all runs, and equally uninformative in all
of them. This is yet another manifestation of the baseline degeneracy:
the total number of spots is not an observable in relative photometry.

There is one final distribution that is instructive to consider: the
distribution of errors on the inferred stellar inclination.
Figure~\ref{fig:calibration_batch_default_inclinations} shows a histogram
of stellar inclination residuals (posterior mean minus true value)
normalized to the posterior standard deviation for all stars across the
100 trials described above. For a correctly calibrated model, this distribution
should equal the standard normal $\mathcal{N}(0, 1^2)$ in the limit of
infinite trials. This, in fact, is roughly what we find (compare to the
orange histogram in the figure). Our posterior has marginally
heavier tails, meaning we tend to \emph{slightly} underestimate the
posterior variance, but in general it is an excellent estimator of
individual stellar inclinations.

@@

Figure~\ref{fig:calibration_hilat}
Figure~\ref{fig:calibration_equatorial}
Figure~\ref{fig:calibration_isotropic}
Figure~\ref{fig:calibration_tinyspots}
Figure~\ref{fig:calibration_default_1}
Figure~\ref{fig:calibration_default_1000}
Figure~\ref{fig:calibration_ld}
Figure~\ref{fig:calibration_ld_1000}
Figure~\ref{fig:calibration_ld_500_no_ld}
Figure~\ref{fig:calibration_unnorm}
Figure~\ref{fig:calibration_bigspots}
Figure~\ref{fig:calibration_variance}
Figure~\ref{fig:calibration_batch_default_1000}

Tiny sunspots are about $70{,}000$ km across, comparable to large sunspots.

@@

@@

\subsection{Prior on sigma}
\label{sec:sigma-prior}
Shrinkage prior on sigma.

\subsection{Miscellaneous}
Inclination distribution is bimodal.

@@

\section{Supplementary Figures}

\begin{figure}[p!]
    \begin{centering}
        \oscaption{calibration_unnorm}{%
            \xxx{TODO: Unnormalized.}
            \label{fig:calibration_unnorm}
        }
    \end{centering}
\end{figure}

\begin{figure}[p!]
    \begin{centering}
        \oscaption{calibration_bigspots}{%
            \xxx{TODO: Large single spot.}
            \label{fig:calibration_bigspots}
        }
    \end{centering}
\end{figure}

\begin{figure}[p!]
    \begin{centering}
        \oscaption{calibration_variance}{%
            \xxx{TODO: Variance in all input parameters.}
            \label{fig:calibration_variance}
        }
    \end{centering}
\end{figure}

\begin{figure}[p!]
    \begin{centering}
        \oscaption{calibration_batch_default_1000}{%
            \xxx{TODO: Batch run with $M=1000$.}
            \label{fig:calibration_batch_default_1000}
        }
    \end{centering}
\end{figure}
